### 2018-07-14 Msc-Slides/main.tex
Slides in progress

\documentclass[compress]{beamer}

%\usepackage[latin1,utf8]{inputenc}

%\usepackage[utf8]{inputenc}
%\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}

\usetheme{shadow}
\usepackage{multirow}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{textpos}
\usepackage{textcomp}
\usepackage{color, colortbl}
\usepackage{fancybox}
\usepackage{boxedminipage}
\usepackage{xcolor}
\usepackage{lipsum}
\usepackage{comment}
\usepackage{rotating}
\usepackage{xcolor}

\usepackage{epsfig}
\usepackage{amsmath,amssymb}
\usepackage{amsthm}
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{boxedminipage}
\usepackage{latexsym}
\usepackage{subfigure}
\usepackage{wrapfig}
\usepackage{algorithm} 
\usepackage{appendixnumberbeamer}

\usepackage{framed}

\usetikzlibrary{shapes}
\usetikzlibrary{arrows}
\usetikzlibrary{automata}
\usepackage{algorithm}
\usepackage[noend]{algorithmic}
\usepackage{pifont}
\usepackage{fancybox}
\usepackage{listings}
\usepackage{changepage}

% Daniele
\usepackage{adjustbox}

\lstset{escapeinside={<@}{@>}}

% needed if you use oopdf2eps script :)
\epstopdfsetup{suffix=}

\beamertemplatenavigationsymbolsempty

\title[BlueTracer:
a Robust API Tracer for Evasive Malware]{BlueTracer: \\
a Robust API Tracer for Evasive Malware}

\author[Simone Nicchi]
{
    \textbf{\textcolor{sapienza}{Simone Nicchi}\\~\\{\vspace{-0.5em}\em\textmd{Thesis Advisor: Prof. Camil Demetrescu \\ Thesis Co-Advisors: Dr. Daniele Cono D'Elia, Dr. Emilio Coppa} }\\~\\{Master of Science in Engineering in Computer Science}}\vspace{-2em}
}

%\pgfdeclareimage[height=1.6cm]{logo}{Immagini/sapienza}
%\logo{\centering \pgfuseimage{logo}}

%\institute[Sapienza Universit\`{a} di Roma, Italy]
%{
  %\inst{1}%
  %Sapienza Universit\`{a} di Roma, Italy\\
  %\and
  %\vskip-2mm
  %\inst{2}%
  %Sapienza Universit\`{a} di Roma, Italy\\
%  \texttt{\space crisafulli.giu@gmail.com}
%}

%\date{\today}
\date{July 20, 2018}
\newcommand{\argmax}{\operatornamewithlimits{argmax}\limits}
\definecolor{forest}{RGB}{21,166,64}
\newcommand\mybox[2][]{\tikz[overlay]\node[ draw=forest, fill=blue!20,inner sep=2pt, anchor=text, rectangle, rounded corners=1mm, #1] {#2};\phantom{#2}}

\begin{document}

\definecolor{sapienza}{rgb}{0.50,0.14,0.20}
\definecolor{LightCyan}{rgb}{0.88,1,1}
\definecolor{Gray}{gray}{0.9}

\newcommand\darrow{\fontsize{36pt}{36pt}\selectfont{\begin{turn}{-90}\bf {\textcolor{sapienza}{\ding{224}}}\end{turn}}}
\newcommand\larrow{\fontsize{36pt}{36pt}\selectfont\bf {\textcolor{sapienza}{\ding{224}}}}
\newcommand\blarrow{\fontsize{36pt}{36pt}\selectfont{\begin{turn}{-135}\bf {\textcolor{sapienza}{\ding{224}}}\end{turn}}}

% insert page number
\newcommand*\oldmacro{}
\let\oldmacro\insertshortauthor% save previous definition
\renewcommand*\insertshortauthor{%
  \leftskip=.3cm% before the author could be a plus1fill ...
  \insertframenumber\,/\,\inserttotalframenumber\hfill\oldmacro}

\newcommand\emlarrow{%
        {\bf {\textcolor{sapienza}{\ding{224}}}}%
}

\setbeamercolor*{alerted text}{bg=white,fg=sapienza}
\setbeamercolor*{structure}{bg=white,fg=sapienza}

\setbeamercolor*{palette primary}{use=structure,fg=white,bg=sapienza}
\setbeamercolor*{palette secondary}{use=structure,fg=white}
\setbeamercolor*{palette tertiary}{use=structure,fg=white}
\setbeamercolor*{palette quaternary}{fg=black,bg=white}

\setbeamercolor*{sidebar}{use=structure,bg=structure.fg}

\setbeamercolor*{palette sidebar primary}{use=structure}
\setbeamercolor*{palette sidebar secondary}{fg=white}
\setbeamercolor*{palette sidebar tertiary}{use=structure}
\setbeamercolor*{palette sidebar quaternary}{fg=white}

\setbeamercolor*{titlelike}{parent=palette primary}

\setbeamercolor*{separation line}{}
\setbeamercolor*{fine separation line}{}

\setbeamercolor{frametitle}{fg=white,bg=sapienza!100}
\setbeamercolor{frametitle right}{fg=white,bg=sapienza!100}
\setbeamercolor{subsection in head/foot}{bg=sapienza}
\setbeamercolor{title in head/foot}{bg=sapienza}

% insert page number
%\newcommand*\oldmacro{}
%\let\oldmacro\insertshortauthor% save previous definition
%\renewcommand*\insertshortauthor{%
%  \leftskip=.3cm% before the author could be a plus1fill ...
%  \insertframenumber\,/\,\inserttotalframenumber\hfill\oldmacro}

\newcommand{\stt}{\small\tt}
\newcommand{\word}[1]{\fontsize{10.4}{10}\textsf{#1}}        % mention of word
\newcommand{\wordb}[2]{\fontsize{10.4}{10}\textsf{#1}$_#2$}        % mention of word
\newcommand{\wordc}[3]{\fontsize{10.4}{10}\textsf{#1}$_#2^#3$}        % mention of word
\newcommand{\wikipage}[1]{\textsc{#1}}        % mention of wikipage

\newcommand\tab[1][0.5cm]{\hspace*{#1}}

\begin{frame}

 \begin{figure}
     \centering
    \vspace{0.5cm}
     \includegraphics[scale=.40]{image/sapienza.png}\\

 \end{figure}

\titlepage
\end{frame}

\section{Introduction}
\subsection{Malware Analysis}

\begin{frame}
    \frametitle{Malware: an increasingly significant problem }

    \begin{figure}
        \includegraphics[scale=0.55]{image/malware2}
    \end{figure}

\end{frame}

\begin{frame}
    \frametitle{Malware Analysis}

    Two main types:
    \medskip
    \begin{itemize}
        \item \textbf{Static Analysis:}\\
        involves the inspection of the different data and code sections of a binary
        \item \textbf{Dynamic Analysis:}\\
        the malware sample is executed and the actions it performs on the environment are observed
    \end{itemize}
    \vspace{1cm}

         \begin{beamerboxesrounded}[shadow=true]{}
    Dynamic analysis strongly favoured as it allows to dodge \\ code obfuscations and deal with a large number of samples
    \end{beamerboxesrounded}

    \begin{textblock*}{2cm}(9.0cm,-6.5cm)
   \includegraphics[width=1.2cm]{image/search.png}% use the \includegraphics command here
	\end{textblock*}

\end{frame}

\subsection{Function Call Monitoring}
\begin{frame}
    \frametitle{Function call monitoring}

Functions can abstract
implementation details providing
a semantically richer representation of some functionality. \\\bigskip
Example:
    \begin{figure}
    	\vspace{-0.4cm}
        \includegraphics[scale=0.6]{image/fcmonitoring}
    \end{figure}

\begin{beamerboxesrounded}[shadow=true]{}
The abstractions embodied by \textbf{system calls} and \textbf{library calls} \\ can be used to grasp the visible behavior of a malicious sample
\end{beamerboxesrounded}

\end{frame}

\begin{frame}
    \frametitle{Implementation of function call monitoring}

	    \begin{beamerboxesrounded}[shadow=true]{API Hooking}
    The interception of function calls provided by dynamically
linked \\ libraries (DLLs)
    \end{beamerboxesrounded}
    \bigskip
	Three broad categories:

    \begin{itemize}
    \item Binary Rewriting
    \begin{itemize}
    \item Call Redirection
    \item Function Rewriting
    \end{itemize}
    \item Virtual Machine Introspection (VMI)
    \item \textbf{Dynamic Binary Instrumentation (DBI)}
    \end{itemize}

    \begin{textblock*}{2cm}(8.3cm,-3cm)
   \includegraphics[width=2cm]{image/hook.png}% use the \includegraphics command here
	\end{textblock*}

\end{frame}

\begin{frame}[fragile]
    \frametitle{Dynamic Binary Instrumentation (DBI)}

A dynamic binary analysis
technique in which the behaviour of an application is inspected at run-time via the
injection of analysis code. 
\bigskip
\begin{block}{}
\begin{lstlisting}[basicstyle=\ttfamily\large,xleftmargin=50pt]
<@\textcolor{red}{record(arg1)}@> 
retval = libcall(arg1, &arg2)
<@\textcolor{red}{record(retval, *arg2)}@> 
\end{lstlisting}
\end{block}
\bigskip
\bigskip
\textbf{Problem 1}: existing products have limited logging capabilites

\end{frame}

\subsection{Evasion}

\begin{frame}
    \frametitle{The threat posed by evasive malware}

		    \begin{beamerboxesrounded}[shadow=true]{Evasive malware}
Malware that conceals its harmful behaviour when detecting a
hostile environment, such as a well-known sandbox
solution
    \end{beamerboxesrounded}
    \bigskip

    \begin{figure}
    	\vspace{-0.2cm}
    	\hspace*{1cm}
        \includegraphics[width=9cm]{image/evasive.pdf}
    \end{figure}

    \textbf{Problem 2}: API hooking techniques in literature are not coupled with mechanisms to hide their presence from evasive malware

\end{frame}

\section{BlueTracer}

\begin{frame}
    \frametitle{Our solution: BlueTracer}

\begin{beamerboxesrounded}[shadow=true]{}
\textbf{BlueTracer} is a robust library and system call tracer for Windows programs specialized in evasive malware
\end{beamerboxesrounded}

\bigskip

Implementation details:
\begin{itemize}
\item Based on the \textbf{Intel Pin} DBI framework
\item Integrated with the \textbf{BluePill} stealthy execution framework
\item Combines reliable external sources of prototypes information
\end{itemize}

\bigskip

Key features:
\begin{itemize}
\item Undetected tracing of input parameters, output buffers and return values of over 17 000 system calls and library calls
\item Logging of asynchronous events
\item Resolution of named constants
\end{itemize}

\end{frame}

\subsection{Intel Pin}

\begin{frame}
    \frametitle{Why Intel Pin ?}

	Characteristics:
	\begin{itemize}
	\item \textbf{User-friendliness}
	\item \textbf{Portability}
	\item \textbf{Transparency}
	\item \textbf{Efficiency}
	\end{itemize}
	\bigskip
	\textbf{Analysis routines:} embody the code to be inserted during the application's execution\\
	\smallskip
	\textbf{Instrumentation routines:} determine where the analysis code has to be placed\\
	\bigskip
	Different analysis and instrumentation granularities
	\begin{itemize}
	\item Instruction, trace, routine and image
	\end{itemize}

    \begin{textblock*}{3cm}(6cm,-6.5cm)
   \includegraphics[width=1.6cm]{image/pin.png}% use the \includegraphics command here
	\end{textblock*}

\end{frame}

\subsection{BluePill}

\begin{frame}
    \frametitle{Integration with BluePill}

\end{frame}

\subsection{Architecture}

\begin{frame}
    \frametitle{BlueTracer's architecture}

    \begin{figure}
    	\vspace{-0.5cm}
    	%\hspace*{1cm}
        \includegraphics[width=11cm]{image/BlueTracer.pdf}
    \end{figure}

\end{frame}

\section{Evaluation}

\subsection{Al-Khaser}
\begin{frame}
    \frametitle{Al-Khaser}

\end{frame}

\begin{frame}
    \frametitle{Sample tracked evasive check}

\end{frame}

\subsection{Evasive malware samples}
\begin{frame}
    \frametitle{Evasive malware samples}

\end{frame}

\section{Conclusions}

\begin{frame}
    \frametitle{Conclusion and future developments}

\end{frame}

\begin{frame}
\begin{center}
{\fontsize{15}{15}\selectfont \textbf{Thank you for your attention!}}
\end{center}
\end{frame}

\end{document}