### 2018-07-17 Msc-Slides/main.tex
Slides re-worked once again

\documentclass[compress]{beamer}

%\usepackage[latin1,utf8]{inputenc}

%\usepackage[utf8]{inputenc}
%\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}

\usetheme{shadow}
\usepackage{multirow}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{textpos}
\usepackage{textcomp}
\usepackage{color, colortbl}
\usepackage{fancybox}
\usepackage{boxedminipage}
\usepackage{xcolor}
\usepackage{lipsum}
\usepackage{comment}
\usepackage{rotating}
\usepackage{xcolor}

\usepackage{epsfig}
\usepackage{amsmath,amssymb}
\usepackage{amsthm}
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{boxedminipage}
\usepackage{latexsym}
\usepackage{subfigure}
\usepackage{wrapfig}
\usepackage{algorithm} 
\usepackage{appendixnumberbeamer}

\usepackage{framed}

\usetikzlibrary{shapes}
\usetikzlibrary{arrows}
\usetikzlibrary{automata}
\usepackage{algorithm}
\usepackage[noend]{algorithmic}
\usepackage{pifont}
\usepackage{fancybox}
\usepackage{listings}
\usepackage{changepage}

% Daniele
\usepackage{adjustbox}

\lstset{escapeinside={<@}{@>}}

% needed if you use oopdf2eps script :)
\epstopdfsetup{suffix=}

\beamertemplatenavigationsymbolsempty

\title[BlueTracer:
a Robust API Tracer for Evasive Malware]{BlueTracer: \\
a Robust API Tracer for Evasive Malware}

\author[Simone Nicchi]
{
    \textbf{\textcolor{sapienza}{Simone Nicchi}\\~\\{\vspace{-0.5em}\em\textmd{Thesis Advisor: Prof. Camil Demetrescu \\ Thesis Co-Advisors: Dr. Daniele Cono D'Elia, Dr. Emilio Coppa} }\\~\\{Master of Science in Engineering in Computer Science}}\vspace{-2em}
}

%\pgfdeclareimage[height=1.6cm]{logo}{Immagini/sapienza}
%\logo{\centering \pgfuseimage{logo}}

%\institute[Sapienza Universit\`{a} di Roma, Italy]
%{
  %\inst{1}%
  %Sapienza Universit\`{a} di Roma, Italy\\
  %\and
  %\vskip-2mm
  %\inst{2}%
  %Sapienza Universit\`{a} di Roma, Italy\\
%  \texttt{\space crisafulli.giu@gmail.com}
%}

%\date{\today}
\date{July 20, 2018}
\newcommand{\argmax}{\operatornamewithlimits{argmax}\limits}
\definecolor{forest}{RGB}{21,166,64}
\newcommand\mybox[2][]{\tikz[overlay]\node[ draw=forest, fill=blue!20,inner sep=2pt, anchor=text, rectangle, rounded corners=1mm, #1] {#2};\phantom{#2}}

\begin{document}

\definecolor{sapienza}{rgb}{0.50,0.14,0.20}
\definecolor{LightCyan}{rgb}{0.88,1,1}
\definecolor{Gray}{gray}{0.9}

\newcommand\darrow{\fontsize{36pt}{36pt}\selectfont{\begin{turn}{-90}\bf {\textcolor{sapienza}{\ding{224}}}\end{turn}}}
\newcommand\larrow{\fontsize{36pt}{36pt}\selectfont\bf {\textcolor{sapienza}{\ding{224}}}}
\newcommand\blarrow{\fontsize{36pt}{36pt}\selectfont{\begin{turn}{-135}\bf {\textcolor{sapienza}{\ding{224}}}\end{turn}}}

% insert page number
\newcommand*\oldmacro{}
\let\oldmacro\insertshortauthor% save previous definition
\renewcommand*\insertshortauthor{%
  \leftskip=.3cm% before the author could be a plus1fill ...
  \insertframenumber\,/\,\inserttotalframenumber\hfill\oldmacro}

\newcommand\emlarrow{%
        {\bf {\textcolor{sapienza}{\ding{224}}}}%
}

\setbeamercolor*{alerted text}{bg=white,fg=sapienza}
\setbeamercolor*{structure}{bg=white,fg=sapienza}

\setbeamercolor*{palette primary}{use=structure,fg=white,bg=sapienza}
\setbeamercolor*{palette secondary}{use=structure,fg=white}
\setbeamercolor*{palette tertiary}{use=structure,fg=white}
\setbeamercolor*{palette quaternary}{fg=black,bg=white}

\setbeamercolor*{sidebar}{use=structure,bg=structure.fg}

\setbeamercolor*{palette sidebar primary}{use=structure}
\setbeamercolor*{palette sidebar secondary}{fg=white}
\setbeamercolor*{palette sidebar tertiary}{use=structure}
\setbeamercolor*{palette sidebar quaternary}{fg=white}

\setbeamercolor*{titlelike}{parent=palette primary}

\setbeamercolor*{separation line}{}
\setbeamercolor*{fine separation line}{}

\setbeamercolor{frametitle}{fg=white,bg=sapienza!100}
\setbeamercolor{frametitle right}{fg=white,bg=sapienza!100}
\setbeamercolor{subsection in head/foot}{bg=sapienza}
\setbeamercolor{title in head/foot}{bg=sapienza}

% insert page number
%\newcommand*\oldmacro{}
%\let\oldmacro\insertshortauthor% save previous definition
%\renewcommand*\insertshortauthor{%
%  \leftskip=.3cm% before the author could be a plus1fill ...
%  \insertframenumber\,/\,\inserttotalframenumber\hfill\oldmacro}

\newcommand{\stt}{\small\tt}
\newcommand{\word}[1]{\fontsize{10.4}{10}\textsf{#1}}        % mention of word
\newcommand{\wordb}[2]{\fontsize{10.4}{10}\textsf{#1}$_#2$}        % mention of word
\newcommand{\wordc}[3]{\fontsize{10.4}{10}\textsf{#1}$_#2^#3$}        % mention of word
\newcommand{\wikipage}[1]{\textsc{#1}}        % mention of wikipage

\newcommand\tab[1][0.5cm]{\hspace*{#1}}

\begin{frame}

 \begin{figure}
     \centering
    \vspace{0.5cm}
     \includegraphics[scale=.40]{image/sapienza.png}\\

 \end{figure}

\titlepage
\end{frame}

\section{Intro.}

\subsection{Malware Analysis}

\iffalse
\begin{frame}
    \frametitle{Malware: an increasingly significant problem }

    \begin{figure}
        \includegraphics[scale=0.55]{image/malware2}
    \end{figure}

\end{frame}
\fi

\begin{frame}
    \frametitle{Malware Analysis}

	Malware is an ever-growing threat

    %Two main types:
    \medskip
    \begin{itemize}
        \item \textcolor{sapienza}{\textbf{Static Analysis:}}\\
        involves the inspection of the different data and code sections of a binary
        \item \textcolor{sapienza}{\textbf{Dynamic Analysis:}}\\
        the malware sample is executed and the actions it performs on the environment are observed
    \end{itemize}
    \vspace{0.5cm}

         \begin{beamerboxesrounded}[shadow=true]{}
    Dynamic analysis strongly favoured as it allows to dodge most code obfuscations and deal with a large number of samples
    \end{beamerboxesrounded}

    \begin{textblock*}{2cm}(8.0cm,-6.1cm)
   \includegraphics[width=1.2cm]{image/search.png}% use the \includegraphics command here
	\end{textblock*}

\end{frame}

\subsection{Function Call Monitoring}
\begin{frame}[fragile]
    \frametitle{Function call monitoring}

\begin{itemize}
\item Functions can abstract
implementation details providing
a semantically richer representation of some functionality
\item The abstractions embodied by \textbf{system calls} and \textbf{library calls} \\ can be used to grasp the visible behavior of a malicious sample 
\end{itemize}
\medskip
\textcolor{sapienza}{\textbf{Example}}:\\
\bigskip

\begin{beamerboxesrounded}[shadow=true]{}

\lstset{
    language=C++,
    stringstyle=\color{sapienza}, % string color
    basicstyle=\linespread{1.5}\ttfamily\small,
}

\begin{lstlisting}[xleftmargin=0pt]
RegCreateKey("...\CurrentVersion\Run\monitor") 
CreateDirectory("C:\Windows\utils")
CreateFile("C\Windows\utils\GFypmMVqJQOEQqy.exe")
\end{lstlisting}

\end{beamerboxesrounded}

\end{frame}

\iffalse
\begin{frame}[fragile]
    \frametitle{Implementing function call monitoring}

	    \begin{beamerboxesrounded}[shadow=true]{API Hooking}
    The interception of function calls provided by dynamically
linked \\ libraries (DLLs)
    \end{beamerboxesrounded}
    \bigskip

%	Three broad categories:

%    \begin{itemize}
%    \item Binary Rewriting
%    \begin{itemize}
%    \item Call Redirection
%    \item Function Rewriting
%    \end{itemize}
%    \item Virtual Machine Introspection (VMI)
%    \item \textbf{Dynamic Binary Instrumentation (DBI)}
%    \end{itemize}

%    \begin{textblock*}{2cm}(8.3cm,-3cm)
%   \includegraphics[width=2cm]{image/hook.png}% use the \includegraphics command here
%	\end{textblock*}

One technique is \textbf{Dynamic Binary Instrumentation (DBI)}.
\begin{itemize}
\item[$\rightarrow$] The behaviour of an application is inspected at run time, without the need of recompiling it - by injecting analysis code
\end{itemize}
\smallskip

\begin{block}{}
\begin{lstlisting}[basicstyle=\ttfamily\large,xleftmargin=50pt]
<@\textcolor{red}{record\_before(libcall\_name, arg1)}@> 
retval = libcall(arg1, &arg2)
<@\textcolor{red}{record\_after(retval, *arg2)}@> 
\end{lstlisting}
\end{block}

\end{frame}
\fi

\iffalse
\begin{frame}[fragile]
    \frametitle{Dynamic Binary Instrumentation (DBI)}

A dynamic binary analysis
technique in which the behaviour of an application is inspected at run-time, without the need of recompiling it, via the
injection of analysis code. 
\bigskip
\begin{block}{}
\begin{lstlisting}[basicstyle=\ttfamily\large,xleftmargin=50pt]
<@\textcolor{red}{record\_before(libcall\_name, arg1)}@> 
retval = libcall(arg1, &arg2)
<@\textcolor{red}{record\_after(retval, *arg2)}@> 
\end{lstlisting}
\end{block}
\bigskip
\textcolor{sapienza}{\textbf{Problems:}}
\begin{enumerate}
\item Existing products have limited logging capabilites
\item API hooking techniques in literature are not coupled with mechanisms to hide their presence from evasive malware
\end{enumerate}

\end{frame}
\fi

\subsection{Problems}
\begin{frame}[fragile]
    \frametitle{Problem 1: limited logging capabilities}

Available API tracing tools have \textbf{limited logging capabilities}
%\begin{itemize}
%\item[--] The amount of recorded information is too little
%\end{itemize}
\medskip
	\begin{block}{}
\begin{lstlisting}[basicstyle=\ttfamily\large,xleftmargin=50pt]
<@\textcolor{red}{record\_before(libcall\_name, arg1)}@> 
retval = libcall(arg1, &arg2)
<@\textcolor{red}{record\_after(retval, *arg2)}@> 
\end{lstlisting}
\end{block}	
\medskip
\begin{itemize}
\item Access to a source of calls related information is required
\begin{itemize}
\item[--] Prototype (number of arguments, data types, input/output)
%\item[--] Extraction
\end{itemize}
\medskip
\item \textbf{Challenge:} heterogeneity of Windows libraries used in malware and lack of well-structured documentation for their prototypes
\end{itemize}

\end{frame}

\begin{frame}[fragile]
    \frametitle{Problem 2: logging only calls made by sample}

It is hard to distinguish the calls made directly by the sample from the ones made within libraries
\begin{itemize}
\item Resulting logs are large and full of irrelevant information
\end{itemize}

    \begin{figure}
    	\vspace{-0.2cm}
    	\hspace*{-1cm}
        \includegraphics[width=12cm]{image/MainImage1.pdf}
    \end{figure}

\end{frame}

\begin{frame}[fragile]
    \frametitle{Problem 3: evasive malware}

     \begin{beamerboxesrounded}[shadow=true]{Evasive malware}
Malware that conceals its harmful behaviour when detecting a
hostile environment, such as a well-known sandbox
solution
    \end{beamerboxesrounded}
    \bigskip
    \bigskip
    \begin{figure}
    	\vspace{-0.8cm}
    	\hspace*{0.5cm}
        \includegraphics[width=10cm]{image/evasive.pdf}
    \end{figure}
\vspace{-0.3cm}
Current tracing techniques are \textbf{easily detectable} and are not coupled with mechanisms to hide their presence

\end{frame}

\iffalse
\subsection{Evasion}
\begin{frame}
    \frametitle{The threat posed by evasive malware}

		    \begin{beamerboxesrounded}[shadow=true]{Evasive malware}
Malware that conceals its harmful behaviour when detecting a
hostile environment, such as a well-known sandbox
solution
    \end{beamerboxesrounded}
    \bigskip

    \begin{figure}
    	\vspace{-0.6cm}
    	\hspace*{0.5cm}
        \includegraphics[width=12cm]{image/evasive.pdf}
    \end{figure}

\end{frame}
\fi

\iffalse
\begin{frame}
    \frametitle{Existing problems}

\section{Problem}	
Two main problems to address:
\\\bigskip
\begin{enumerate}
\item Available API tracing tools have \textbf{limited logging capabilities}
\smallskip
\begin{itemize}
\item[--] The amount of recorded information is too little
\item[--] It is hard to distinguish the calls made directly by the sample from the ones made within libraries
\end{itemize}
\bigskip
\item Current API hooking techniques are \textbf{easily detectable} and are not coupled with mechanisms to hide their presence
from evasive malware
\end{enumerate}

\end{frame}
\fi

\section{BlueTracer}

\subsection{Overview}
\begin{frame}
    \frametitle{BlueTracer: accurate API logging}
    \setbeamerfont*{itemize/enumerate subbody}{parent=itemize/enumerate body}

\medskip

\begin{beamerboxesrounded}[shadow=true]{}
\textbf{BlueTracer} is a robust library and system call tracer for Windows programs specialized in evasive malware based on Intel Pin
\end{beamerboxesrounded}

\bigskip

\textcolor{sapienza}{\textbf{The tool possesses a remarkable logging power:}}
\begin{itemize}
%\item Integration of reliable external sources \\ (\textbf{Dr. %Memory}
%and \textbf{CISCO PyREBox})
%\item Logging operations are challenging 
%by the heterogeneity of Windows libraries used in malware and the lack of well-structured documentation for their prototypes
%\begin{itemize}
\item Integration of reliable external sources \\ (\textbf{Dr. Memory}
and \textbf{CISCO PyREBox})
%\end{itemize} 
%\end{itemize}

%\textcolor{sapienza}{\textbf{Key features:}}
%\begin{itemize}
\item Stealthy tracing of input parameters, output buffers and return values of over 17,000 system calls and library calls
\item Logging of asynchronous events
%\item Allows to trace only calls from the main executable
\end{itemize}

\end{frame}

\begin{frame}
    \frametitle{BlueTracer: robust against evasive malware}
\setbeamerfont*{itemize/enumerate subbody}{parent=itemize/enumerate body}

\textcolor{sapienza}{\textbf{Solution to the detection problem:}}
\begin{itemize}
\item Seamless integration with the \textbf{BluePill} stealthy execution framework
%\item Combines reliable external sources of prototypes information
\end{itemize}

%\textcolor{sapienza}{\textbf{Building blocks:}}
%\begin{itemize}
%\item Based on the \textbf{Intel Pin} DBI framework
%\item Integrated with the \textbf{BluePill} stealthy execution framework
%\item Combines reliable external sources of prototypes information
%\end{itemize}

\bigskip
\bigskip
	    \begin{figure}
    	\vspace{-0.8cm}
    	%\hspace*{1cm}
        \includegraphics[width=8cm]{image/BluePill.pdf}
    \end{figure}

%\textcolor{sapienza}{\textbf{Faced implementation challenges:}}
%\begin{itemize}
%\item Optimal use of
%Intel Pin's capabilities with respect to run-time CPU and memory costs
%\item Addressing some inherent limitations of Intel Pin
%\end{itemize}

\end{frame}

\begin{frame}
    \frametitle{BlueTracer: focussed tracing of the sample's actions}
\setbeamerfont*{itemize/enumerate subbody}{parent=itemize/enumerate body}

Only calls made directly by the sample are recorded through the use of context-sensitive introspection
\begin{itemize}
\item Stack pointer inspection
\end{itemize}    
\medskip
	    \begin{figure}
    	\vspace{-0.3cm}
    	%\hspace*{1cm}
        \includegraphics[width=10cm]{image/MainImageTree.pdf}
    \end{figure}

\end{frame}

\iffalse
\subsection{Intel Pin}

\begin{frame}
    \frametitle{Why Intel Pin ?}

	Characteristics:
	\begin{itemize}
	\item \textbf{User-friendliness}
	\item \textbf{Portability}
	\item \textbf{Transparency}
	\item \textbf{Efficiency}
	\item \textbf{Robustness}
	\end{itemize}
	\medskip
	\textbf{Analysis routines:} embody the code to be inserted during the application's execution\\
	\smallskip
	\textbf{Instrumentation routines:} determine where the analysis code has to be placed\\
	\bigskip

%	Different analysis and instrumentation granularities
%	\begin{itemize}
%	\item Instruction, trace, routine and image
%	\end{itemize}

    \begin{textblock*}{3cm}(6cm,-6.5cm)
   \includegraphics[width=1.6cm]{image/pin.png}% use the \includegraphics command here
	\end{textblock*}

\end{frame}

\subsection{BluePill}

\begin{frame}
    \frametitle{Integration with BluePill}
	\textbf{BluePill} is a software toolkit which:
	\begin{itemize}
	\item Allows the simulation of a real production environment a specific malware
sample was intended for
	\item Conceals any virtualization artifacts and software setup which might set off
evasion
	\end{itemize}
	\medskip

	    \begin{figure}
    	\vspace{-0.3cm}
    	%\hspace*{1cm}
        \includegraphics[width=7.5cm]{image/BluePill.pdf}
    \end{figure}

\end{frame}
\fi

\subsection{Architecture}

\begin{frame}
    \frametitle{BlueTracer's architecture}

    \begin{figure}
    	\vspace{-0.5cm}
    	%\hspace*{1cm}
        \includegraphics[width=11cm]{image/BlueTracer.pdf}
    \end{figure}

\end{frame}

\section{Validation}

\subsection{Al-Khaser}
\begin{frame}
    \frametitle{Validation (1) with Al-Khaser}
    \begin{beamerboxesrounded}[shadow=true]{}
	\textbf{Al-Khaser} is an open-source application which performs common checks employed by malware families to determine if they are being
executed in an analysis environment.
\end{beamerboxesrounded}
\medskip
Checks divided in categories:
\begin{itemize}
\item \textbf{Anti-Debugging}
\item \textbf{Timing-based}
\item \textbf{Human Interaction Detection}
\item \textbf{Anti-Virtualization}
\item \textbf{Detection of Analysis Tools}
\end{itemize}
\medskip
\begin{beamerboxesrounded}[shadow=true]{}
BlueTracer: 
\begin{itemize}
\item remained undetected thanks to its integration with BluePill
\item managed to track all the checks
\end{itemize}
\end{beamerboxesrounded}

    \begin{textblock*}{3cm}(7.5cm,-5.5cm)
   \includegraphics[width=2cm]{image/lights.png}% use the \includegraphics command here
	\end{textblock*}

\end{frame}

\subsection{Evasive malware samples}
\begin{frame}
    \frametitle{Validation (2) with highly evasive malware samples}

	Five highly evasive samples collected by Joe Security:

    \begin{table}[h]
\vspace*{-0.1cm}
\begin{center}
\resizebox{9.2cm}{!}{
 \begin{tabular}{||c c c||}
 \hline
 \textbf{ID} & \textbf{MD5} & \textbf{Name} \\ 
 \hline\hline
 1 & 0af4ef5069f47a371a0caf22ae2006a6 & \textit{banker} \\ 
 \hline
 2 & 9437eabf2fe5d32101e3fbf9f6027880 & \textit{dropper} \\
 \hline
 3 & cbdda646a20d95f078393506ecdc0796 & \textit{trojan}\\
 \hline
 4 & cfdd16225e67471f5ef54cab9b3a5558 & Olympic\\
 \hline
 5 & ef694b89ad7addb9a16bb6f26f1efaf7 & CCleaner\\ 
 \hline
\end{tabular}
}
\end{center}
\end{table}
\vspace{-0.1cm}
Evaluation was done manually and is a time-consuming process:
\begin{itemize}
\item Check if logs are congruous with Joe Security reports
\item Process Monitor as ground truth for system activity
\end{itemize}
\medskip
\begin{beamerboxesrounded}[shadow=true]{}
The
logs collected by BlueTracer reveal behaviors consistent with the analysis reports
authored by Joe Security
\end{beamerboxesrounded}

\end{frame}

\begin{frame}[fragile]
    \frametitle{Example of tracked malevolent action}

\begin{beamerboxesrounded}[shadow=true]{}
Tracing a particular action of a malware instance allows to understand in detail what the sample's intentions are
\end{beamerboxesrounded}
\bigskip    
\textbf{Example}: dropping a malicious executable
\vspace*{0.3cm}

    %\begin{figure}
    %	\vspace{-0.5cm}
    	%\hspace*{1cm}
    %    \includegraphics[width=11cm]{image/SampleCall}
    %\end{figure}

	\lstset{
    language=C++,
    frame=tb, % draw a frame at the top and bottom of the code block
    tabsize=4, % tab space width
    showstringspaces=false, % don't mark spaces in strings
    numbers=none, % display line numbers on the left
    commentstyle=\color{ao}, % comment color
    %keywordstyle=\color{blue}, % keyword color
    stringstyle=\color{red}, % string color
    basicstyle=\linespread{1.3}\footnotesize\ttfamily,
    basewidth = {.48em},
   escapeinside={<@}{@>}
}

\begin{lstlisting}
~~1116~~ 138 kernel32.dll!CopyFileA
138 	arg 0: c:\Users\Simuset\Desktop\sample1.exe 
(name=lpExistingFileName, type=char*, size=0x1)
138 	arg 1: <@\alert{C:\textbackslash Windows\textbackslash system32\textbackslash $\dagger$\textbackslash ffpb6966.exe}@> 
(name=lpNewFileName, type=char*, size=0x1)
138 	arg 2: 0x0 (name=bFailIfExists, type=(long/int), size=0x4)
138    executed kernel32.dll!CopyFileA =>
138 	retval: 0x1 (name=Return value, type=(long/int), size=0x4)
\end{lstlisting}

\iffalse
\begin{lstlisting}
~~972~~ 1397 kernel32.dll!CreateFileW
1397 	arg 0: <@\alert{C:\textbackslash Windows  GFypmMVqJQOEQqy.exe}@> 
(name=lpFileName, type=wchar_t*, size=0x2)
1397 	arg 1: 0x40000000 (name=dwDesiredAccess, type=DWORD, size=0x4)
1397 	arg 2: 0x1 (name=dwShareMode, type=DWORD, size=0x4)
1397 	arg 3: 0x00000000 (name=lpSecurityAttributes, type=struct*|_SECURITY_ATTRIBUTES, size=0x60)
1397 	arg 4: 0x2 
(name=dwCreationDisposition, type=DWORD, size=0x4)
1397 	arg 5: 0x0 (name=dwFlagsAndAttributes, type=DWORD, size=0x4)
1397 	arg 6: 0x0 (name=hTemplateFile, type=DWORD, size=0x4)
1397    executed kernel32.dll!CreateFileW =>
1397 	retval: 0x128 (name=Return value, type=DWORD, size=0x4))
\end{lstlisting}
\fi

\lstset{
    language=C++,
    frame=tb, % draw a frame at the top and bottom of the code block
    tabsize=4, % tab space width
    showstringspaces=false, % don't mark spaces in strings
    numbers=left, % display line numbers on the left
    commentstyle=\color{ao}, % comment color
    keywordstyle=\color{blue}, % keyword color
    stringstyle=\color{red}, % string color
    basicstyle=\footnotesize\ttfamily,
    basewidth = {.48em}
}

\end{frame}

\section{Conclusions}

\begin{frame}
    \frametitle{Conclusions}

	\textbf{\textcolor{sapienza}{Contribution:}} \\
	\smallskip	
	\tab Design and implementation of \textbf{BlueTracer},
a robust library and system call tracer for Windows programs specialized in evasive
malware.
	\\\bigskip
	\textbf{\textcolor{sapienza}{Future Developments:}} \\
	\begin{itemize}
	\item Automatic methodology for large-scale evaluation 
	\item Improvement of logging capabilities
	\item Usage of log filtering and aggregation techniques
	\end{itemize}

\end{frame}

\begin{frame}
\begin{center}
{\fontsize{15}{15}\selectfont \textbf{Thank you for your attention!}}
\end{center}
\end{frame}

\end{document}